---
title: "Week 4 Data Analysis"
author: "Alexander Jones"
date: "`Week 4`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reshape2)
```


## Overview

This report present preliminary data analysis for Week 3 Example Data. Specifically, this will report basic descriptive statistics as well as report tests of the following predictions:

1. People will feel more guilt when:
  + 1.1 They cause the shock (vs. their partner)
  + 1.2 Their partner was morally good (vs. bad)
  + 1.3 The shock was stronger (high > medium > low)
2. Women will report more guilt than men
3. Agreeable people will report more guilt
4. Neurotic people will report more guilt

and 

5. People will be more generous when:
  +5.1 They caused the shock (vs. their partner)
  +5.2 Their partner was morally good (vs. bad)
  +5.3 The shock was stronger (high > medium > low)
6. Women will be more generous than men
7. Agreeable people will be more generous
8. Neurotic people will be more generous


```{r load}

#Load the data
data<-read.csv("Week 3/Processed Data/Week 3 PROCESSED.csv")


#Recode Variables:
data$PIN<-as.factor(data$PIN)
data$sex<-as.factor(data$sex)
data$shock<-as.factor(data$shock)
data$shockCause<-as.factor(data$shockCause)
data$pMoral<-as.factor(data$pMoral)

```



## Descriptives

Descriptive Statistics for guilt:
```{r descGuilt}

c(summary(data$guilt),"SD"=sd(data$guilt,na.rm=T))

```


Descriptive Statistics for generosity:
```{r descGenerosity}

#Generosity before the shock
c(summary(data$preShare),"SD"=sd(data$preShare,na.rm=T))

#Generosity after the shock
c(summary(data$postShare),"SD"=sd(data$postShare,na.rm=T))
```


Descriptive for personality:
```{rdescPers}

#Descriptive for Agreeableness
c(summary(data$agree),"SD"=sd(data$agree,na.rm=T))

#Descriptive for Neuroticism 
c(summary(data$neuro),"SD"=sd(data$neuro,na.rm=T))

```




## Guilt Analyses

### Prediction 1.1

Prediction 1.1 was supported. People felt more guilty when they caused the shock than when their partner did:
```{r p1.1}

#Run the t-test for prediction 1.1
p1.1<-t.test(data$guilt~data$shockCause)

#Output the result
p1.1

```


### Prediction 1.2

Prediction 1.2 was supported. People felt more guilty when the shock was to a morally good person than a morally bad person:
```{r p1.2}

#Run the t-test for prediction 1.2
p1.2<-t.test(data$guilt~data$pMoral)

#Output the result
p1.2

```


### Prediction 1.3

Prediction 1.3 was supported. People felt more guilty when their partner recieved a stronger shock:
```{r p1.3}

#Run the One-Way ANOVA 1.3
p1.3<-lm(guilt~shock,data=data)
  
#Output the result
anova(p1.3)

#Output the group means
tapply(data$guilt,data$shock,function(x) mean (x,na.rm=T))

```


### Prediction 2
Prediction 2 was not supported. Women did not report more guilt than men.
```{r p2}

#Run the t-test
p2<-t.test(guilt~sex, data=data)
  
#Output the result
p2
```


### Prediction 3
Prediction 3 was supported. Agreeable people feel slightly more guilty than others.
```{r p3}

#Run the correlation test
p3<-cor.test(data$guilt, data$agree)

#Output the result
p3
```


### Prediction 4
Prediction 4 was not supported. There was not a correlation between neurotic people and guilt.
```{r}

#Run the correlation test
p4<-cor.test(data$guilt, data$neuro)

#Output the result
p4
```



## Generosity Analyses

```{r melt}

#Melt our data from wide format to long format
dataMelt<-melt(data,
               measure.vars=c("preShare","postShare"),
               variable.name=c("prePost"),
               value.name=c("share")
               )

```

### Prediction 5.1
Prediction was supported. People who caused the shock themselves were more generous than those who did not cause the shock and were *less* generous when the partner caused the shock:
```{r p5.1}

#Run the 2 x 2 ANOVA (Within Subjects)
p5.1<-aov(share~prePost*shockCause+Error(PIN),data=dataMelt)

#Output the result
summary(p5.1)

#Output the means
tapply(dataMelt$share,dataMelt[,c(24,21)],function(x)mean(x,na.rm=T))
       
```


### Prediction 5.2

```{r p5.2}

#Run the 2 x 2 ANOVA
p5.2<-aov(share~prePost*pMoral+Error(PIN),data=dataMelt)
  
#Output the result
summary(p5.2)

#Output the means
tapply(dataMelt$share,dataMelt[,c(24,22)],function(x)mean(x,na.rm=T))

```


